---
title: "Getting Started with rq"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with rq}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`rq` is a simple, script-basked workflow manager. Specifically, it excels in
combining simple workflows (say, filling some data from source `X, Y and Z`, 
aggregating it and then saving to a database) into DAGs, and then executing
whilst honoring pre-requisites.

This is a user guide which aims to cover all of the functionalities that are
core to `rq`.

### Creating nodes

`JobNode`s are the building blocks of a workflow in `rq`. They describe a 
single unit of work; a single `R` script, alongside a `name` which describes
the workflow.

In this user guide, we will make some temp files to simulate scripts that 
must run.

```{r}
library(rq)

dummy_script <- function(...) {
  path <- tempfile(fileext = ".R")
  writeLines(c(...), path)
  path
}

print_script <- dummy_script("print('Hello, world');")
print_job <- JobNode$new(
  'Job which prints hello, world!', 
  print_script, 
  upstream=list(), 
  on_fail = 'halt'
)
```

We have just created our first `JobNode`. `rq` is not fussed about whether you use
these nodes as standalone workflows; if your workflow uses one script, and for whatever
reason, you'd like to abstract the running of this script - you are free to do so.

```{r}
if (print_job$is_ready()) {
  handle <- print_job$run()
  handle$wait()
  handle$read_all_output_lines()
}
```

The anatomy of a job is described by two major components:
1. Is the job ready? This is decided by it's `upstream` components - things that it relies on to be finished
2. Running the job - the job should only be ran once it is ready to run

### Orchestrating and visualising workflows

Having a single script is great, but how do you create workflows? Workflows are
created by aggregating multiple `JobNode`s into the `JobOrchestrator`. Nothing
fancy is required - just the creation of a bunch of `JobNode`s with appropiate 
(and unique) names, and then feed them into the orchestrator.

The `JobOrchestrator` will check for you whether your workflow is a compliant one
(it must be directed and acyclic!).

```{r}
sleep_then_print_script <- dummy_script(
  "Sys.sleep(2)",
  "print('God - we are awfully tired')",
  "Sys.sleep(3)",
  "print('Wake up!')"
)

wake_up_script <- dummy_script(
  "print('Alright, we are all woken up!')"
)

sleep_then_print_job <- JobNode$new(
  "Lazy job",
  script_path = sleep_then_print_script,
  upstream = list()
)

wake_up_job <- JobNode$new(
  "Motivated job",
  script_path = wake_up_script,
  upstream = list(sleep_then_print_job)
)
```

We've made two job nodes - the job for 'waking up' and being 'motivated' relies on the sleepy job - we declare this using the `upstream` input. We then feed these jobs into the `JobOrchestrator`.

```{r}
jo <- JobOrchestrator$new()
jo$add_job(sleep_then_print_job)
jo$add_job(wake_up_job)
```

Before we run anything, we might want to visualise how our workflow runs.
We can this using the `visualise`.

```{r}
jo$visualise()
```

`visualise` gives a interactive view of how your jobs are going/went.

We can then get to actually running the `JobOrchestrator`. We want to first check that 

```{r}
jo
jo$run()
```

Cool! `Motivated job` waited until `Lazy job` was finished before running successfully. If we `visualise` again...

```{r}
jo$visualise()
```

We can see that both jobs have succeeded.

### A more realistic example

Let's say you are a quantitative research working on a pipeline which pulls
data from `Bloomberg`, aggregates the data, and then writes the aggregations
into a database. We've just added a new data source from `FRED`, which is
aggregated in the same aggregation script.

Let's set up that workflow.

```{r}
bloomberg_data <- dummy_script(
  "print('Pulling data from Bloomberg...')",
  "Sys.sleep(5)",
  "print('Success!')"
)

fred_data <- dummy_script(
  "print('Pulling data from FRED...')",
  "Sys.sleep(2)",
  "q(status = 1)"
)

agg_script <- dummy_script(
  "print('Aggregating...')",
  "Sys.sleep(3)"
)

db_write <- dummy_script(
  "print('Saving to db...')",
  "Sys.sleep(1)"
)

bbg_job <- JobNode$new('BBG Data', bloomberg_data)
fred_job <- JobNode$new('FRED Data', fred_data, on_fail="halt")
agg_job <- JobNode$new('Aggregation', agg_script, upstream = list(bbg_job, fred_job))
db_job <- JobNode$new('Write to DB', db_write, upstream = list(agg_job))

jo <- JobOrchestrator$new()
jo$add_job(bbg_job)
jo$add_job(fred_job)
jo$add_job(agg_job)
jo$add_job(db_job)

jo$visualise()
```

You will notice that the FRED data script fails. Let's see what happens when we run the orchestrator.

```{r}
jo$run()
```

We can also visualise where the workflow failed.

```{r}
jo$visualise()
```

So the orchestrator halts on `FRED Data`, as we've told it to halt. In data science and ETL workflows, it's often the 
case (particularly for things like scraped data) that things don't work deterministically. We can instead, add
`retries`.

```{r}
jo$remove_job('FRED Data')

# fails 80% of the time
fred_data <- dummy_script(
  "print('Pulling data from FRED...')",
  "Sys.sleep(2)",
  "if (runif(1) < 0.8) q(status = 1)"
)

fred_job <- JobNode$new('FRED Data', fred_data)

jo$add_job(fred_job, retries=10)

# relink the aggregation job to the fred job
jo$add_upstream('Aggregation', 'FRED Data')

# reset the orchestrator
jo$reset_all()

jo$visualise()
```

Okay, let's try running this again.

```{r}
jo$run()
```

The FRED data retried once, and then succeeded.

```{r}
jo$visualise()
```