---
title: "Getting Started with rq"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with rq}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`rq` is a simple, script-basked workflow manager. Specifically, it excels in
combining simple workflows (say, filling some data from source `X, Y and Z`, 
aggregating it and then saving to a database) into DAGs, and then executing
whilst honoring pre-requisites.

This is a user guide which aims to cover all of the functionalities that are
core to `rq`.

### Creating nodes

`JobNode`s are the building blocks of a workflow in `rq`. They describe a 
single unit of work; a single `R` script, alongside a `name` which describes
the workflow.

In this user guide, we will make some temp files to simulate scripts that 
must run.

```{r}
library(rq)

dummy_script <- function(...) {
  path <- tempfile(fileext = ".R")
  writeLines(c(...), path)
  path
}

print_script <- dummy_script("print('Hello, world');")
print_job <- JobNode$new(
  'Job which prints hello, world!', 
  print_script, 
  upstream=list(), 
  on_fail = 'halt'
)
```

We have just created our first `JobNode`. `rq` is not fussed about whether you use
these nodes as standalone workflows; if your workflow uses one script, and for whatever
reason, you'd like to abstract the running of this script - you are free to do so.

```{r}
if (print_job$is_ready()) {
  handle <- print_job$run()
  handle$wait()
  handle$read_all_output_lines()
}
```

The anatomy of a job is described by two major components:
1. Is the job ready? This is decided by it's `upstream` components - things that it relies on to be finished
2. Running the job - the job should only be ran once it is ready to run

### Orchestrating and visualising workflows

Having a single script is great, but how do you create workflows? Workflows are
created by aggregating multiple `JobNode`s into the `JobOrchestrator`. Nothing
fancy is required - just the creation of a bunch of `JobNode`s with appropiate 
(and unique) names, and then feed them into the orchestrator.

The `JobOrchestrator` will check for you whether your workflow is a compliant one
(it must be directed and acyclic!).

```{r}

sleep_then_print_script <- dummy_script(
  "Sys.sleep(2)",
  "print('God - we are awfully tired')",
  "Sys.sleep(3)",
  "print('Wake up!')"
)

wake_up_script <- dummy_script(
  "print('Alright, we are all woken up!')"
)

sleep_then_print_job <- JobNode$new(
  "Lazy job",
  script_path = sleep_then_print_script,
  upstream = list()
)

wake_up_job <- JobNode$new(
  "Motivated job",
  script_path = wake_up_script,
  upstream = list(sleep_then_print_job)
)
```

We've made two job nodes - the job for 'waking up' and being 'motivated' relies on the sleepy job - we declare this using the `upstream` input. We then feed these jobs into the `JobOrchestrator`.

```{r}
jo <- JobOrchestrator$new()
jo$add_job(sleep_then_print_job)
jo$add_job(wake_up_job)
```

Before we run anything, we might want to visualise how our workflow runs.
We can this using the `WorkflowPrinter`.

```{r}
wp <- WorkflowPrinter$new()
wp$print_workflow(jo)
```

We can then get to actually running the `JobOrchestrator`.

```{r}
jo$run()
```

Cool! `Motivated job` waited until `Lazy job` was finished before running successfully.
